Demes: a specification for interchanging complex demography

### 2022-04-07 paper.tex

\krtcomment{We need to revisit the paragraph(s) below.  We do not assume individuals are exchangeable, because we do not
	require that all individuals have the same fitness. See, e.g., p 60 of Wakeley for why fitness == exchangeability
	and why that is required for the standard coalescent.}

Within a population, individuals are assumed to be exchangeable, but there are
also parameters for nonrandom mating, such as selfing or cloning rates, which
give the probability that offspring are generated from one generation to the
next by self-fertilisation or cloning of an individual. Cloning and selfing
rates can take any value between zero and one, and if both are specified then
the selfing rate is the probability an offspring is generated through
self-fertilisation given it was produced sexually (i.e. with probability one
minus the cloning rate). We omit any details regarding selection.
Thus, the history of population size changes does not depend on mean fitnesses within
a population nor on differences in mean fitness between populations.
This indepedence of population size and
selection is often called ``soft" selection \citep{christiansen1975hard}.

>>>>

\krtcomment{We need to revisit the paragraph(s) below.  We do not assume individuals are exchangeable, because we do not
	require that all individuals have the same fitness. See, e.g., p 60 of Wakeley for why fitness == exchangeability
	and why that is required for the standard coalescent.}
\mhcomment{The `exchangeability' assumption also breaks down under 
	different reproductive modes, since the sampling of the alleles affects coalescent times. 
	For example, in a highly selfing species the coalescent time for two different alleles 
	within an individual is much lower than two alleles from different individuals. I've tried 
	to rewrite the below paragraph to update and clarify the reproduction parameters in 
	line with the discussions on Github. Finally, 'rates' imply events occur as Poisson-like 
	processes over continuous time, whereas the values used here instead define the 
	\emph{fraction} of reproductions of each type. Shall we rename to selfing/clonal `fractions',
	 or would that mess up the rest of the schema too much?}

Each population has a selfing rate and cloning rate assigned to it, where each denotes 
the probability that offspring are generated from one generation to the next by either 
self-fertilisation or cloning of an individual. More specifically, denote the clonal rate by 
$\sigma$ and the selfing rate by $S$. In each population a proportion of offspring 
$\sigma$ reproduce clonally, while $1-\sigma$ reproduce sexually. Within the 
sexually--reproducing offspring, $S$ were born via self--fertilisation while the rest were 
outcrossed. Hence $S$ and $\sigma$ can take any value between zero and one, but 
can sum to more than one. See \citet{hartfield2016facsexcoal} for more details of the 
mathematical properties of these rates in a coalescent context. \mhcomment{If requested 
I can quickly draw up a schematic diagram to illustrate these partitions.}

We omit any details regarding selection.
Thus, the history of population size changes does not depend on mean fitnesses within
a population nor on differences in mean fitness between populations.
This indepedence of population size and
selection is often called ``soft" selection \citep{christiansen1975hard}.
